Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2014-07-17 cordillera-cycle.tex
Polish section 4.2.2.

Pehaps one of the most persistent landscape feature is the Skeena Mountains ice
cap, which persisted throughout the entire last glacial cycle
(Figs.~\ref{fig:snapshots} and~\ref{fig:duration}). Regardless of the applied
forcing (Fig.~\ref{fig:snapshots}, middle panels), this ice cap appears to
survive MIS~3, and serves as a nucleation centre at the onset of the glacial
readvance towards the LGM (MIS~2). The importance of residual ice for Noeth
American glacial history leading up to the LGM has been illustrated by the
MIS~3 residual ice bodies in northern and eastern Canada as nucleation centres
for a much more extensive MIS~2 configuration \citep{Kleman.etal.2010}.

The presence of a Skeena Mountains ice cap during most of the last glacial
cycle can be explained by a more widespread\renote{
    I mean widespread, precipitation there is lower than elsewhere.},
current winter precipitation for that region there than for other parts of the
modelling domain (Fig.~\ref{fig:atm}). Along most of the north-western coast of
North America, coastal mountain ranges form a pronounced topographic barrier
for westerly winds, capturing atmospheric moisture in the form of orographic
precipitation, and resulting in arid interior lowlands. However, near the
centre of our modelling domain, this barrier is less pronounced than elsewhere,
allowing westerly winds to carry moisture\renote{
    I do not agree with ``carry precipitation''. When reading this I get the
    picture of water drops falling diagonaly under action of the wind. I do not
    deny that transport of precipitation happens and
    is an important process for understanding orographic
    precipitation (think of why it is raining so much in the Swedish mountains
    even when the wind comes from the west), but it is not what I am trying to
    describe here. I wrote ``moisture'' and this is what I mean.}
further inland, until it is progressively captured by the extensive, but mildly
elevated group of the Skeena Mountains in north-central British Columbia.

>>>>

It is generally believed that the Cordilleran ice sheet formed by the
coalescence of of several mountain-centred ice caps \citep{Davis.Mathews.1944}.
In our simulations, major ice-dispersal centres, visible on the modelled ice
cover duration maps (Fig.~\ref{fig:duration}), are located over the Coast
Mountains (CM), the Columbia and Rocky mountains (CR), the Skeena Mountains
(SM), and the Selwyn and MacKenzie Mountains (SMK). Although the Coast,
Columbia, Rocky and Skeena Mountains (CM, CR, SM) are covered by mountain
glaciers for most of the model time, providing durable nucleation centre for
the initiation of ice sheet growth, this is not the case for the Selwyn and
MacKenzie mountains (SMK), where ice cover on the highest peaks is limited to a
fraction of the last glacial cycle. In other words, the Selwyn and MacKenzie
mountains appear as a secondary ice-dispersal centre which ``activates'' only
during the coldest periods of the last glacial cycle.

Perhaps the most striking feature displayed by the distributions of modelled
ice cover is the persistence of the Skeena Mountains ice cap throughout the
entire last glacial cycle and its predominance over other ice-dispersal centres
(Figs.~\ref{fig:snapshots} and~\ref{fig:duration}). Regardless of the applied
forcing, this ice cap appears to survive MIS~3 (Fig.~\ref{fig:snapshots},
middle panels), and serves as a nucleation centre at the onset of the glacial
readvance towards the LGM (MIS~2). The importance of residual ice for North
American glacial history leading up to the LGM has been illustrated by the
MIS~3 residual ice bodies in northern and eastern Canada as nucleation centres
for a much more extensive MIS~2 configuration \citep{Kleman.etal.2010}\renote{
    Does this sentence really fit here? I feel that this sudden mention of the
    Laurentide ice sheet perturbates a bit the flow. Moreover, knowing that
    the Laurentide ice sheet is probably the main driver for the 100\,ka
    sea-level cycle, it would have been surprising if it had completely
    vanished at MIS~3. But for the Cordilleran ice sheet, usually thought of
    as fast-responding and intermittent, I feel that the presence of residual
    ice is less trivial.}.

The presence of a Skeena Mountains ice cap during most of the last glacial
cycle can be explained by a currently more widespread\renote{
    I mean widespread, precipitation there is lower than elsewhere.}
winter precipitation for that region there than for other parts of the
modelling domain (Fig.~\ref{fig:atm}). Along most of the north-western coast of
North America, coastal mountain ranges form a pronounced topographic barrier
for westerly winds, capturing atmospheric moisture in the form of orographic
precipitation, and resulting in arid interior lowlands. However, near the
centre of our modelling domain, this barrier is less pronounced than elsewhere,
allowing westerly winds to carry moisture\renote{
    I do not agree with ``carry precipitation''. When reading this I get the
    picture of water drops falling diagonaly under action of the wind. I do not
    deny that transport of precipitation happens and
    is an important process for understanding orographic
    precipitation (think of why it is raining so much in the Swedish mountains
    even when the wind comes from the west), but it is not what I am trying to
    describe here. I wrote ``moisture'' and this is what I mean.}
further inland, until it is progressively captured by the extensive, but mildly
elevated group of the Skeena Mountains in north-central British Columbia.
